---
title: "cort analysis"
author: "mns"
date: "March 1, 2016"
output: pdf_document
---
This document combines the corticosterone data from the individual EIAs into one large dataset. That dataset is used to investigate the effect of the treatments on the corticosterone level of individual frogs. Corticosterone was measured in the water at the beginning of the 48 hour exposure and at the end of the exposure. A standard curve was created for each EIA plate individually as abundance units vs. log (concentration) values. 

Import files used:
```{r, echo=FALSE}
setwd("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final")
#import sample identification key
#key_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_022916.csv")
#import cort values for plate1 02/22/16
cort_plate5_022216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate5_022216.csv")
plate1<-cort_plate5_022216
#import cort values for plate2 02/23/16
cort_plate6_022316 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate6_022316.csv")
plate2<-cort_plate6_022316
#import cort values for plate3 02/26/16
cort_plate14_022616 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate14_022616.csv")
plate3<-cort_plate14_022616
#import cort values for plate4 02/29/16
cort_plate7_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate7_022916.csv")
plate4<-cort_plate7_022916
#import cort values for plate5 03/02/16
cort_plate5_030216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate5_030216.csv")
plate5<-cort_plate5_030216
#import key for plate 1
EIA_Cort_022216_412_KEY <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022216_412_KEY.csv")
key1<-EIA_Cort_022216_412_KEY
#import key for plate 2
EIA_Cort_022316_412_key <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022316_412_key.csv")
key2<-EIA_Cort_022316_412_key
#import key for plate 3
EIA_Cort_022616_key <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022616_key.csv")
key3<-EIA_Cort_022616_key
#import key for plate 4
key_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_022916.csv")
key4<-key_022916
#import key for plate 5
key_030216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_030216.csv")
key5<-key_030216
```

Average duplicate values for samples and merge

```{r, echo=FALSE}
library(plyr)
#### plate 1 ####
#convert to lower case
plate1$cell<-tolower(plate1$cell)
plate1a<-merge(plate1, key1, by.x='cell', by.y='plate_key')
View(plate1a)
#get mean of sample duplicate values
plate1.mean<-ddply(plate1a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate1b<-merge(key1, plate1.mean, by.x='sample_name', by.y='sample_name')
View(plate1b)
#remove duplicate rows
plate1c<-plate1b[!duplicated(plate1b$sample_name),]
plate1c$plate<-'plate1'
View(plate1c)

#### plate 2 ####
#convert to lower case
plate2$cell<-tolower(plate2$cell)
plate2a<-merge(plate2, key2, by.x='cell', by.y='plate_key')
View(plate2a)
#get mean of sample duplicate values
plate2.mean<-ddply(plate2a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate2b<-merge(key2, plate2.mean, by.x='sample_name', by.y='sample_name')
View(plate2b)
#remove duplicate rows
plate2c<-plate2b[!duplicated(plate2b$sample_name),]
plate2c$plate<-'plate2'
View(plate2c)

#### plate 3 ####
#convert to lower case
plate3$cell<-tolower(plate3$cell)
plate3a<-merge(plate3, key3, by.x='cell', by.y='plate_key')
View(plate3a)
#get mean of sample duplicate values
plate3.mean<-ddply(plate3a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate3b<-merge(key3, plate3.mean, by.x='sample_name', by.y='sample_name')
View(plate3b)
#remove duplicate rows
plate3c<-plate3b[!duplicated(plate3b$sample_name),]
plate3c$plate<-'plate3'
View(plate3c)

#### plate 4 ####
#convert to lower case
plate4$cell<-tolower(plate4$cell)
plate4a<-merge(plate4, key4, by.x='cell', by.y='plate_key')
View(plate4a)
#get mean of sample duplicate values
plate4.mean<-ddply(plate4a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate4b<-merge(key4, plate4.mean, by.x='sample_name', by.y='sample_name')
View(plate4b)
#remove duplicate rows
plate4c<-plate4b[!duplicated(plate4b$sample_name),]
plate4c$plate<-'plate4'
View(plate4c)

#### plate 5 ####
#convert to lower case
plate5$cell<-tolower(plate5$cell)
plate5a<-merge(plate5, key5, by.x='cell', by.y='plate_key')
View(plate5a)
#get mean of sample duplicate values
plate5.mean<-ddply(plate5a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate5b<-merge(key5, plate5.mean, by.x='sample_name', by.y='sample_name')
View(plate5b)
#remove duplicate rows
plate5c<-plate5b[!duplicated(plate5b$sample_name),]
plate5c$plate<-'plate5'
View(plate5c)

#### combine dataframes from each plate into 1 df #### 
# convert numeric to factor for experiment column
plate2c$experiment<-as.factor(as.character(plate2c$experiment))
plate3c$experiment<-as.factor(as.character(plate3c$experiment))
plate4c$experiment<-as.factor(as.character(plate4c$experiment))
plate5c$experiment<-as.factor(as.character(plate5c$experiment))
#combine all plates into 1 df
cort_all<-rbind(plate1c, plate2c, plate3c, plate4c, plate5c)
View(cort_all)
# sort by sample number
cort_all_sort<-cort_all[order(cort_all$sample_no),]
#sum samples with multiple cort cells
c49<-(1.284533e+03+2.092193e+01) 
c54<-(5.855812e+01+4.860801e+00)
s26<-(2.235718e+02+6.634500e+00)
s15<-(30.35141825+1.357286e+02)
#add1<-c(as.factor(as.character("exp4_C49base")),as.factor(as.character("a11g10")),as.factor(as.character("4")), as.factor(as.character("1")), "C49", as.numeric(c49))
cort_all_sort$sample_name<-as.factor(as.character(cort_all_sort$sample_name))
#cort_all_sort1<-rbind(cort_all_sort, add1)
write.csv(cort_all_sort, "cort_out.csv")
```

Anova for experimental day effects
```{r}
cort_out2 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_out3.csv")
View(cort_out2)
cort1<-cort_out2
cort1$experiment<-as.factor(as.character(cort1$experiment))
summary(cort1)
anova1<-lm(V1~experiment, data=subset(cort1, time=="1"))
summary(anova1)
anova2<-lm(V1~experiment, data=subset(cort1, time=="2"))
summary(anova2)

boxplot(V1~experiment, data=subset(cort1, time=="2"), main="Cort End")
boxplot(V1~experiment, data=subset(cort1, time=="1"), main= "Cort Beginning")

#create dataframe separate for initial and final time periods
cort1.base<-cort1[cort1$time=="1",]
View(cort1.base)
cort1.end<-cort1[cort1$time=="2",]
View(cort1.end)

#merge based on experiment number and sample name
cort2<-merge(cort1.base, cort1.end, by=c("sample_no", "experiment"))
summary(cort2)
View(cort2)
dim(cort2)
#subtract final measurement from initial cort measurement
cort2$cort.diff<-cort2$V1.y-cort2$V1.x

anova3<-lm(cort.diff~experiment, data=cort2)
summary(anova3)
boxplot(cort.diff~experiment, data=cort2, main="Cort difference")
```
Treatment 1= control
Treatment 2= pesticide
Treatment 3= bullfrog
Treatment 4= bullfrog+pesticide

Anova and boxplots for treatment effects:
```{r}
sample_list <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/sample_list.csv")
cort3<-merge(cort1, sample_list, by=c("sample_no", "experiment"))
View(cort3)
cort3$treatment<-as.factor(as.character(cort3$treatment))
cort3$time<-as.factor(as.character(cort3$time))
#anovas for effect of treatment on cort level at beginning and end
m1<-lm(V1~treatment, data=subset(cort3, time=="1"))
summary(m1)
m2<-lm(V1~treatment, data=subset(cort3, time=="2"))
summary(m2)

boxplot(V1~treatment, data=subset(cort3, time=="2"), xlab="treatment", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="End measurement")
boxplot(V1~treatment, data=subset(cort3, time=="1"), xlab="treatment", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="Base measurement")

#create dataframe separate for initial and final time periods
cort3.base<-cort3[cort3$time=="1",]
View(cort3.base)
cort3.end<-cort3[cort3$time=="2",]
View(cort3.end)
cort4<-merge(cort3.base, cort3.end, by=c("sample_no", "experiment"))
#cort4<-merge(cort3, sample_list, by=c("sample_no", "experiment"))
cort4$cort.diff<-cort4$V1.y-cort4$V1.x
View(cort4)

#anova for treatment effect on cort difference
m3<-lm(cort.diff~treatment.x, data=cort4)
summary(m3)

boxplot(cort.diff~treatment, data=cort4, xlab="treatment.x", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="difference measurement")
#anova for EIA plate effect on cort difference
model.plate.beg<-lm(V1~plate, data=subset(cort3, time=="1"))
summary(model.plate.beg)
model.plate.end<-lm(V1~plate, data=subset(cort3, time=="2"))
summary(model.plate.end)

```
Linear models with more than one predictor factor
```{r}
m3<-lm(V1.y~treatment.x+plate.x, data=cort4)
summary(m3)

model.plate.end2<-lm(V1~treatment+plate, data=subset(cort3, time=="2"))
summary(model.plate.end2)
```

When comparing the experiment time period, the significant differences are in the comparison of cort levels to experimental time period is that the beginning base line cort level between experiment 3 and 4. For end time period intercept and experimental time 2 is significantly different. When comparing treatment levels to corticosterone levels, Treatment 2 and 4 is significantly different in end time period Cort levels. In difference cort levels or beginning cort levels there is no signficant treatment effects. Treatment 2 is carbaryl only. Treatment 4 is carbaryl and bullfrog predation stress.   

ANOVAS without plate 4 for cort beginning and end time periods:
```{r}
#create dataset without plate4
cort5<-cort3[cort3$plate!="plate4",]
summary(cort5)
#anova without plate4 for end time period
model.end.woplate4<-lm(V1~treatment+plate, data=subset(cort5, time=="2"))
summary(model.end.woplate4)
model.end.woplate5<-lm(V1~Pesticide+predator+plate, data=subset(cort5, time=="2"))
summary(model.end.woplate5)

boxplot(V1~treatment, data=subset(cort5, time=="2"))
model.end.woplate4.2<-lm(V1~treatment, data=subset(cort5, time=="2"))
summary(model.end.woplate4.2)
model.end.woplate4.3<-lm(V1~plate, data=subset(cort5, time=="2"))
summary(model.end.woplate4.3)
#anova without plate4 for base time period
model.end.woplate4<-lm(V1~treatment+plate, data=subset(cort5, time=="1"))
summary(model.end.woplate4)
boxplot(V1~treatment, data=subset(cort5, time=="1"))
model.end.woplate4.2<-lm(V1~treatment, data=subset(cort5, time=="1"))
summary(model.end.woplate4.2)

```


Exploratory plots:
```{r}
plot(cort4$V1.x, cort4$cort.diff)
abline(0, 0)
plot(cort4$V1.y, cort4$cort.diff)
abline(0,0)
summary(cort4$cort.diff)
summary(cort4$V1.x)
summary(cort4$V1.y)
#make id for experiment sample number to compare base and end measurements for same frog 
cort4$id <- with(cort4, interaction(experiment,  sample_no))
plot(cort4$V1.x, cort4$V1.y)
#plot base and end cort measurements for same frog on same plot
plot(cort4$V1.y~cort4$id, pch=1, col="red" )
points(cort4$V1.x~cort4$id, col="red")
```
Re-calibrated plate4 standard curve with average slope and intercept from other plates. Reimporting and re-analyzing data.
###################################################################
Import files used:
```{r, echo=FALSE}
setwd("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final")
#import sample identification key
#key_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_022916.csv")
#import cort values for plate1 02/22/16
cort_plate5_022216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate5_022216.csv")
plate1<-cort_plate5_022216
#import cort values for plate2 02/23/16
cort_plate6_022316 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate6_022316.csv")
plate2<-cort_plate6_022316
#import cort values for plate3 02/26/16
cort_plate14_022616 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate14_022616.csv")
plate3<-cort_plate14_022616
#import cort values for plate4 02/29/16
cort_plate7_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate7_022916_2.csv")
plate4<-cort_plate7_022916
#import cort values for plate5 03/02/16
cort_plate5_030216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_plate5_030216.csv")
plate5<-cort_plate5_030216
#import key for plate 1
EIA_Cort_022216_412_KEY <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022216_412_KEY.csv")
key1<-EIA_Cort_022216_412_KEY
#import key for plate 2
EIA_Cort_022316_412_key <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022316_412_key.csv")
key2<-EIA_Cort_022316_412_key
#import key for plate 3
EIA_Cort_022616_key <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/EIA_Cort_022616_key.csv")
key3<-EIA_Cort_022616_key
#import key for plate 4
key_022916 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_022916.csv")
key4<-key_022916
#import key for plate 5
key_030216 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/key_030216.csv")
key5<-key_030216
```

Average duplicate values for samples and merge

```{r, echo=FALSE}
library(plyr)
#### plate 1 ####
#convert to lower case
plate1$cell<-tolower(plate1$cell)
plate1a<-merge(plate1, key1, by.x='cell', by.y='plate_key')
View(plate1a)
#get mean of sample duplicate values
plate1.mean<-ddply(plate1a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate1b<-merge(key1, plate1.mean, by.x='sample_name', by.y='sample_name')
View(plate1b)
#remove duplicate rows
plate1c<-plate1b[!duplicated(plate1b$sample_name),]
plate1c$plate<-'plate1'
View(plate1c)

#### plate 2 ####
#convert to lower case
plate2$cell<-tolower(plate2$cell)
plate2a<-merge(plate2, key2, by.x='cell', by.y='plate_key')
View(plate2a)
#get mean of sample duplicate values
plate2.mean<-ddply(plate2a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate2b<-merge(key2, plate2.mean, by.x='sample_name', by.y='sample_name')
View(plate2b)
#remove duplicate rows
plate2c<-plate2b[!duplicated(plate2b$sample_name),]
plate2c$plate<-'plate2'
View(plate2c)

#### plate 3 ####
#convert to lower case
plate3$cell<-tolower(plate3$cell)
plate3a<-merge(plate3, key3, by.x='cell', by.y='plate_key')
View(plate3a)
#get mean of sample duplicate values
plate3.mean<-ddply(plate3a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate3b<-merge(key3, plate3.mean, by.x='sample_name', by.y='sample_name')
View(plate3b)
#remove duplicate rows
plate3c<-plate3b[!duplicated(plate3b$sample_name),]
plate3c$plate<-'plate3'
View(plate3c)

#### plate 4 ####
#convert to lower case
plate4$cell<-tolower(plate4$cell)
plate4a<-merge(plate4, key4, by.x='cell', by.y='plate_key')
View(plate4a)
#get mean of sample duplicate values
plate4.mean<-ddply(plate4a, .(sample_name), function(x) mean(x$pg.ml))
#merge df with means with original plate key
plate4b<-merge(key4, plate4.mean, by.x='sample_name', by.y='sample_name')
View(plate4b)
#remove duplicate rows
plate4c<-plate4b[!duplicated(plate4b$sample_name),]
plate4c$plate<-'plate4'
View(plate4c)

#### plate 5 ####
#convert to lower case
plate5$cell<-tolower(plate5$cell)
plate5a<-merge(plate5, key5, by.x='cell', by.y='plate_key')
View(plate5a)
#get mean of sample duplicate values
plate5.mean<-ddply(plate5a, .(sample_name), function(x) mean(x$cort.pg.ml))
#merge df with means with original plate key
plate5b<-merge(key5, plate5.mean, by.x='sample_name', by.y='sample_name')
View(plate5b)
#remove duplicate rows
plate5c<-plate5b[!duplicated(plate5b$sample_name),]
plate5c$plate<-'plate5'
View(plate5c)

#### combine dataframes from each plate into 1 df #### 
# convert numeric to factor for experiment column
plate2c$experiment<-as.factor(as.character(plate2c$experiment))
plate3c$experiment<-as.factor(as.character(plate3c$experiment))
plate4c$experiment<-as.factor(as.character(plate4c$experiment))
plate5c$experiment<-as.factor(as.character(plate5c$experiment))
#combine all plates into 1 df
cort_all<-rbind(plate1c, plate2c, plate3c, plate4c, plate5c)
View(cort_all)
# sort by sample number
cort_all_sort<-cort_all[order(cort_all$sample_no),]
#sum samples with multiple cort cells
c49<-(1.284533e+03+2.092193e+01) 
c54<-(5.855812e+01+4.860801e+00)
s26<-(2.235718e+02+6.634500e+00)
s15<-(30.35141825+1.357286e+02)
#add1<-c(as.factor(as.character("exp4_C49base")),as.factor(as.character("a11g10")),as.factor(as.character("4")), as.factor(as.character("1")), "C49", as.numeric(c49))
cort_all_sort$sample_name<-as.factor(as.character(cort_all_sort$sample_name))
#cort_all_sort1<-rbind(cort_all_sort, add1)
write.csv(cort_all_sort, "cort_out.csv")
```


Anova for experimental day effects
```{r}
cort_out4 <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/CORT/final/cort_out4.csv")
View(cort_out4)
scort1<-cort_out4
scort1$experiment<-as.factor(as.character(scort1$experiment))
summary(scort1)
anova1<-lm(V1~experiment, data=subset(scort1, time=="1"))
summary(anova1)
anova2<-lm(V1~experiment, data=subset(scort1, time=="2"))
summary(anova2)

boxplot(V1~experiment, data=subset(scort1, time=="2"), main="Cort End")
boxplot(V1~experiment, data=subset(scort1, time=="1"), main= "Cort Beginning")

#create dataframe separate for initial and final time periods
scort1.base<-scort1[scort1$time=="1",]
View(scort1.base)
scort1.end<-scort1[scort1$time=="2",]
View(scort1.end)

#merge based on experiment number and sample name
scort2<-merge(scort1.base, scort1.end, by=c("sample_no", "experiment"))
summary(scort2)
View(scort2)
dim(scort2)
#subtract final measurement from initial cort measurement
scort2$cort.diff<-scort2$V1.y-scort2$V1.x

anova3<-lm(cort.diff~experiment, data=scort2)
summary(anova3)
boxplot(cort.diff~experiment, data=scort2, main="Cort difference")
```

```{r}
sample_list <- read.csv("~/Dropbox/amphib_metabolomics/PIP/data/sample_list.csv")
scort3<-merge(scort1, sample_list, by=c("sample_no", "experiment"))
View(scort3)
scort3$treatment<-as.factor(as.character(scort3$treatment))
scort3$time<-as.factor(as.character(scort3$time))
#anovas for effect of treatment on cort level at beginning and end
m1<-lm(V1~treatment, data=subset(scort3, time=="1"))
summary(m1)
m2<-lm(V1~treatment, data=subset(scort3, time=="2"))
summary(m2)

boxplot(V1~treatment, data=subset(scort3, time=="2"), xlab="treatment", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="End measurement")
boxplot(V1~treatment, data=subset(scort3, time=="1"), xlab="treatment", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="Base measurement")
```
Merge dfs together to create cort difference column
```{r}
#create dataframe separate for initial and final time periods
scort3.base<-scort3[scort3$time=="1",]
View(scort3.base)
scort3.end<-scort3[scort3$time=="2",]
View(scort3.end)
scort4<-merge(scort3.base, scort3.end, by=c("sample_no", "experiment"))
#cort4<-merge(cort3, sample_list, by=c("sample_no", "experiment"))
scort4$cort.diff<-scort4$V1.y-scort4$V1.x
View(scort4)

#anova for treatment effect on cort difference
m3<-lm(cort.diff~treatment.x, data=scort4)
summary(m3)

boxplot(cort.diff~treatment.x, data=scort4, xlab="treatment.x", names=c("control","pesticide","predator","pesticide+predator"), ylab="Cort pg/mL", main="difference measurement")
```

```{r}
#cort begin time
m4<-lm(V1.x~treatment.x+experiment, data=scort4)
summary(m4)

#cort end time
m4<-lm(V1.y~treatment.x+experiment, data=scort4)
summary(m4)
m4<-lm(V1.y~treatment.x+plate.x+experiment, data=scort4)
summary(m4)
m5<-lm(V1.y~Pesticide.x+predator.x+experiment+plate.x, data=scort4)
summary(m5)
m5<-lm(V1.y~experiment, data=scort4)
summary(m5)
m7<-lm(V1.y~Pesticide.x*predator.x+experiment, data=scort4)
summary(m7)
boxplot(V1.y~Pesticide.x+predator.x, data=scort4, main="Cort final")
#cort difference 
m3<-lm(cort.diff~Pesticide.x+predator.x+experiment, data=scort4)
summary(m3)
m3<-lm(cort.diff~treatment.x+experiment, data=scort4)
summary(m3)
boxplot(cort.diff~Pesticide.x+predator.x, data=scort4, main="Cort diff")
```

